---
title: "2022 MCLC Survey Results"
author: "The CSG Justice Center"
date: "4/7/2023"
output:
  html_document
---

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
#######################################
# PROJECT: More Community, Less Confinement (2022)
# PURPOSE: Impute values, national estimates, state estimates and costs
# AUTHOR: JSM
#######################################

###################################################
#OVERALL SETUP
#load necessary packages
library(openxlsx)
library(dplyr)
library(readr)
library(reshape)
library(readxl)
library(tidyverse)
library(data.table)
library(formattable)
library(scales)
library(mice)
library(VIM)
library(finalfit)
library(janitor)
library(Hmisc)
library(eeptools)
library(csgjcr)
library(scales)
library(xtable)

#read automated_clean to get data, reformat
source("code/survey/01_library_functions.R")
source("code/survey/02_import.R")
source("code/survey/03_clean.R")

##SETUP
#years of data
year     <- c('2018','2019','2020','2021')
yearnum  <- c(1,     2,     3,     4)
refyear  <- c('2019','2020','2021')
refnum   <- 1:3

#number of survey variables
numvar   <- 1:14

#labels
var.labels = c(year                                       = "Year",
               overall_admissions                         = "Overall admissions",
               probation_violation_admissions             = "Total probation violation admissions",
               parole_violation_admissions                = "Total parole violation admissions",
               technical_probation_violation_admissions   = "Admissions for technical violations, probation", 
               technical_parole_violation_admissions      = "Admissions for technical violations, parole",
               new_offense_probation_violation_admissions = "Admissions for new crime violations, probation",
               new_offense_parole_violation_admissions    = "Admissions for new crime violations, parole",
               
               overall_population                         = "Overall population",
               probation_violation_population             = "Total probation violation population",
               parole_violation_population                = "Total parole violation population",
               technical_probation_violation_population   = "Technical violator population, probation",
               technical_parole_violation_population      = "Technical violator population, parole",
               new_offense_probation_violation_population = "New crime violator population, probation",
               new_offense_parole_violation_population    = "New crime violator population, parole"
)

################################################################################
# IMPUTATION
# MICE
# National Estimates w/95% CI's
################################################################################

#calculate imputed values (multiple imputation)
source("code/survey/MImp1.R")

######
#produce confidence intervals
#post multiple imputation t-test
#grab response columns for creating estimated counts/CIs
######

source("code/survey/MImp2.R")

##############################################################
##################CREATE A TABLE HERE!!!!!!!!!!!!#############
#combine
table2018 <- cbind(nat.2018all,natCI.2018all) %>%
  mutate(across(where(is.numeric),formattable::comma,0),
         id     = 1:14,
         ' '    = '95% Confidence Interval',
         '2018' = paste0('(',`1`,', ',`2`,')')
         )
table2019 <- cbind(nat.2019all,natCI.2019all) %>%
  mutate(across(where(is.numeric),formattable::comma,0),
         '2019' = paste0('(',`1`,', ',`2`,')'))
table2020 <- cbind(nat.2020all,natCI.2020all) %>%
  mutate(across(where(is.numeric),formattable::comma,0),
         '2020' = paste0('(',`1`,', ',`2`,')'))
table2021 <- cbind(nat.2021all,natCI.2021all) %>%
  mutate(across(where(is.numeric),formattable::comma,0),
         '2021' = paste0('(',`1`,', ',`2`,')'))

tablevals.pretty <- c('Overall Admissions',
                     'Admissions for Probation Violations',
                     'Admissions for Parole Violations',
                     'Admissions for Technical Probation Violations',
                     'Admissions for Technical Parole Violations',
                     'Admissions for New Crime Probation Violations',
                     'Admissions for New Crime Parole Violations',
                     
                     'Overall Population',
                     'Population of Probation Violators',
                     'Population of Parole Violators',
                     'Population of Technical Probation Violations',
                     'Population of Technical Parole Violations',
                     'Population of New Crime Probation Violations',
                     'Population of New Crime Parole Violations'
                     )

tab.full <- data.table(cbind(table2018$Estimate,table2019$Estimate,table2020$Estimate,table2021$Estimate)) %>%
  mutate(id = 1:14,
         ' ' = 'Estimate', .before="V1") %>%
  dplyr::rename('2018' = "V1",
                '2019' = "V2",
                '2020' = "V3",
                '2021' = "V4") %>%
  rbind(
    data.table(' ' = tablevals.pretty,
               id  = 1:14,
               id2 = 0),
    
    data.table(
      cbind(
      ' '    = table2018$` `,
      '2018' = table2018$`2018`,
      '2019' = table2019$`2019`,
      '2020' = table2020$`2020`,
      '2021' = table2021$`2021`,
      'id'   = table2018$id
      )),
    fill = TRUE) %>%
  mutate(id     = as.numeric(id),
         `2018` =  ifelse(` `=="Estimate",format(as.numeric(`2018`), nsmall=0, digits = 3, scientific = F,big.mark=","),`2018`),
         `2019` =  ifelse(` `=="Estimate",format(as.numeric(`2019`), nsmall=0, digits = 3, scientific = F,big.mark=","),`2019`),
         `2020` =  ifelse(` `=="Estimate",format(as.numeric(`2020`), nsmall=0, digits = 3, scientific = F,big.mark=","),`2020`),
         `2021` =  ifelse(` `=="Estimate",format(as.numeric(`2021`), nsmall=0, digits = 3, scientific = F,big.mark=","),`2021`)
         ) %>%
  arrange(id,id2) %>%
  select(-c(id,id2))

##############################################################
##############################################################

#############################################################################
###############CALCULATING STATE ESTIMATES
#############################################################################

source("code/survey/MImp3.R")

#############################
###COSTS#####################

source("code/survey/Costs.R")

```

## National Estimates

These estimates incorporate BJS total population and admissions data in calculations across all years.

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}

print(xtable(tab.full[1:21,],
                           align = c("llcccc")),
      include.rownames = FALSE,
      type             = "html")

```

<br>
<br>


```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
print(xtable(tab.full[22:42,],
                           align = c("llcccc")),
      include.rownames = FALSE,
      type             = "html")
```

## `r as.numeric(last(refyear))` Costs

The total annual cost for incarcerating **probation violators** in `r as.numeric(last(refyear))` is `r dollar_format()(total.pr.vp)`. The total annual cost for incarcerating **parole violators** in `r as.numeric(last(refyear))` is `r dollar_format()(total.pa.vp)`.

<br>

\noindent The total annual cost for incarcerating just those on **probation with technical violations** in `r as.numeric(last(refyear))` is `r dollar_format()(total.pr.tvp)`. The total annual cost for incarcerating just those on **parole with technical violations** in `r as.numeric(last(refyear))` is `r dollar_format()(total.pa.tvp)`.

<br>

\noindent The total annual cost for incarcerating just those on **probation with new offense violations** in `r as.numeric(last(refyear))` is `r dollar_format()(total.pr.novp)`. The total annual cost for incarcerating just those on **parole with new offense violations** in `r as.numeric(last(refyear))` is `r dollar_format()(total.pa.novp)`.

<br>

### `r as.numeric(last(refyear))` Costs by State

Below is a table of these costs by state (no imputation). Note that missing cells indicate that a state did not submit `r as.numeric(last(refyear))` population data:

<br>

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
print(xtable(cost.final[c("State","cost.vp.state","cost.tvp.state")] %>%
                              mutate(
                                `Community Supervision Violator Cost` = case_when(cost.vp.state >=0 ~ dollar_format()(cost.vp.state),
                                                                                  TRUE ~ NA_character_),
                                `Technical Violator Cost`             = case_when(cost.vp.state >=0 ~ dollar_format()(cost.tvp.state),
                                                                                  TRUE ~ NA_character_)
                                     ) %>%
                              select(-c(cost.vp.state,cost.tvp.state)),
                           align = c("llrr")),
      include.rownames = FALSE,
      type             = "html")
```
