---
title: "2022 MCLC Survey Results"
author: "The CSG Justice Center"
date: "4/13/2023"
output:
  html_document
---

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
#######################################
# PROJECT: More Community, Less Confinement (2022)
# PURPOSE: Impute values, national estimates, state estimates and costs
# AUTHOR: JSM
#######################################

###################################################
#OVERALL SETUP
#load necessary packages
library(openxlsx)
library(dplyr)
library(readr)
library(reshape)
library(readxl)
library(tidyverse)
library(data.table)
library(formattable)
library(scales)
library(mice)
library(VIM)
library(finalfit)
library(janitor)
library(Hmisc)
library(eeptools)
library(csgjcr)
library(scales)
library(xtable)

#for binding blank dataframes
cbind.fill <- function(...){
  nm <- list(...) 
  nm <- lapply(nm, as.matrix)
  n <- max(sapply(nm, nrow)) 
  do.call(cbind, 
          lapply(nm, function (x) 
            rbind(x, matrix( , n-nrow(x), ncol(x)))
            )
          ) 
  }

##CHANGE INPUTS!!!#####################################
#years of data
year     <- as.character(2018:2021)
refyear  <- as.character(2019:2021)

#for labeling tables in output
tablevals.pretty <- c('Overall Admissions',
                     'Admissions for Probation Violations',
                     'Admissions for Parole Violations',
                     'Total Admissions for Technical Violations',
                     'Admissions for Technical Probation Violations',
                     'Admissions for Technical Parole Violations',
                     'Total Admissions for New Crime Violations',
                     'Admissions for New Crime Probation Violations',
                     'Admissions for New Crime Parole Violations',
                     
                     'Overall Population',
                     'Population of Probation Violators',
                     'Population of Parole Violators',
                     'Total Population of Technical Violators',
                     'Population of Technical Probation Violators',
                     'Population of Technical Parole Violators',
                     'Total Population of New Crime Violators',
                     'Population of New Crime Probation Violators',
                     'Population of New Crime Parole Violators'
                     )

################################################################################
#CLEAN & PREP DATA
################################################################################

#read automated_clean to get data, reformat
source("03_clean.R")

#coding missings for calculating national estimates
#There are three types of special missings:
# -99 = REFUSED TO ANSWER SURVEY
##      A value of -99 is assigned for states that left data blank or submitted NAs (did not submit any data)
# -88 = MISSING DATA INTENTIONALLY
##      A value of -88 is assigned for states that submitted 0's, NA's or left data blank *intentionally*
# -77 = DATA DOESN'T EXIST
##      A value of -77 is assigned for states were parole was abolished

adm_pop_analysis_with_bjs1 <- adm_pop_analysis_with_bjs %>%
  mutate(
    #if the lowest level of aggregation is non-missing for both probation and parole (e.g., technical violation admissions probation/parole), then sum to create total (e.g., total technical violation admissions)
    #if the lowest level of aggregation is missing for either or both probation and parole (e.g., technical violation admissions probation/parole), then summed total is MISSING for total (e.g., total technical violation admissions)
    total_technical_violation_admissions   = case_when(
      states %nin% c('Maine','Washington') ~ as.numeric(technical_probation_violation_admissions   + technical_parole_violation_admissions),
      TRUE ~ as.numeric(total_technical_violation_admissions)),
    total_new_offense_violation_admissions = case_when(
      states %nin% c('Maine','Washington') ~ as.numeric(new_offense_probation_violation_admissions + new_offense_parole_violation_admissions),
      TRUE ~ as.numeric(total_new_offense_violation_admissions)),
    total_technical_violation_population   = case_when(
      states %nin% c('Maine','Washington') ~ as.numeric(technical_probation_violation_population   + technical_parole_violation_population),
      TRUE ~ as.numeric(total_technical_violation_population)),
    total_new_offense_violation_population = case_when(
      states %nin% c('Maine','Washington') ~ as.numeric(new_offense_probation_violation_population + new_offense_parole_violation_population),
      TRUE ~ as.numeric(total_new_offense_violation_population)),
    
    ##SPECIAL STATE CASES
    #MAINE (ADMISSIONS ONLY)- all parole variables get 0
    #                         set total new offenses to missing (already conducted in clean_03.R program)
    #                         back calculation new offense probation violations from probation violation total and technical probation violations
    parole_violation_admissions                = case_when(
      states == 'Maine' ~ 0,
      TRUE ~ as.numeric(parole_violation_admissions)),
    technical_parole_violation_admissions      = case_when(
      states == 'Maine' ~ 0,
      TRUE ~ as.numeric(technical_parole_violation_admissions)),
    new_offense_parole_violation_admissions    = case_when(
      states == 'Maine' ~ 0,
      TRUE ~ as.numeric(new_offense_parole_violation_admissions)),
    new_offense_probation_violation_admissions = case_when(
      states == 'Maine' ~ as.numeric(probation_violation_admissions - technical_probation_violation_admissions),
      TRUE ~ as.numeric(new_offense_probation_violation_admissions)),

    #WASHINGTON- all lowest aggregations should be missing
    #            next level aggr. should be used instead of sum
    probation_violation_admissions             = case_when(
      states == 'Washington' ~ NA,
      TRUE ~ probation_violation_admissions),
    parole_violation_admissions                = case_when(
      states == 'Washington' ~ NA,
      TRUE ~ parole_violation_admissions),
    technical_parole_violation_admissions      = case_when(
      states == 'Washington' ~ NA,
      TRUE ~ technical_parole_violation_admissions),
    technical_probation_violation_admissions   = case_when(
      states == 'Washington' ~ NA,
      TRUE ~ technical_probation_violation_admissions),
    new_offense_parole_violation_admissions    = case_when(
      states == 'Washington' ~ NA,
      TRUE ~ new_offense_parole_violation_admissions),
    new_offense_probation_violation_admissions = case_when(
      states == 'Washington' ~ NA,
      TRUE ~ new_offense_probation_violation_admissions),

    probation_violation_population             = case_when(
      states == 'Washington' ~ NA,
      TRUE ~ probation_violation_population),
    parole_violation_population                = case_when(
      states == 'Washington' ~ NA,
      TRUE ~ parole_violation_population),
    technical_parole_violation_population      = case_when(
      states == 'Washington' ~ NA,
      TRUE ~ technical_parole_violation_population),
    technical_probation_violation_population   = case_when(
      states == 'Washington' ~ NA,
      TRUE ~ technical_probation_violation_population),
    new_offense_parole_violation_population    = case_when(
      states == 'Washington' ~ NA,
      TRUE ~ new_offense_parole_violation_population),
    new_offense_probation_violation_population = case_when(
      states == 'Washington' ~ NA,
      TRUE ~ new_offense_probation_violation_population)
    )

################################################################################
# IMPUTATION
# MICE
# National Estimates w/95% CI's
################################################################################
yearnum <- 1:length(year)
refnum  <- 1:(max(yearnum)-1)

#calculate imputed values (multiple imputation)
source("MImp1.R")

######
#produce confidence intervals
#post multiple imputation t-test
#grab response columns for creating estimated counts/CIs
######

source("MImp2.R")

##############################################################
################## CREATE A TABLE HERE #######################
table2018 <- cbind(nat.2018all,natCI.2018all) %>%
  mutate(across(where(is.numeric),formattable::comma,0),
         id     = numvar,
         ' '    = '95% Confidence Interval',
         '2018' = paste0('(',`V1`,', ',`V2`,')')
         )

for (i in 2019:max(year)) {
  temp1 <- get(paste0("nat.",i,"all"))
  temp2 <- get(paste0("natCI.",i,"all"))
  test  <- cbind(temp1,temp2) %>%
    mutate(across(where(is.numeric),formattable::comma,0),
           !!as.name(i) := paste0('(',`V1`,', ',`V2`,')'))
  assign(paste0("table",i),test,envir = .GlobalEnv)
}

#estimates
tab.full1 <- data.table()
for (i in year) {
  temp <- get(paste0("table",i))$Estimate
  tab.full1  <- data.table(cbind(tab.full1,temp))
  setnames(tab.full1, "temp", i)
  tab.full1 <- tab.full1[, lapply(.SD, as.character)]
}

#confidence intervals
CIS <- data.table(' '  = table2018$` `,
                  'id' = table2018$id)
for (i in year) {
  temp <- get(paste0("table",i))[[i]]
  CIS  <- cbind(CIS,temp)
  setnames(CIS, "temp", i)
}

#combine
tab.full <- tab.full1 %>%
  mutate(id = numvar,
         ' ' = 'Estimate') %>%
  relocate(any_of(c('id',' '))) %>%
  rbind(
    data.table(' ' = tablevals.pretty,
               id  = numvar,
               id2 = 0),
    CIS,
    fill = TRUE) %>%
  mutate(id     = as.numeric(id)) %>%
  arrange(id,id2) %>%
  select(-c(id,id2))

#############################################################################
###############CALCULATING STATE ESTIMATES
#############################################################################

source("MImp3.R")

#############################
###COSTS#####################

source("Costs.R")

```

## National Estimates

These estimates incorporate BJS total population and admissions data in calculations across all years.

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}

print(xtable(tab.full[1:((max(numvar)/2)*3),],
                           align = c("llcccc")),
      include.rownames = FALSE,
      type             = "html")

```

<br>
<br>

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
print(xtable(tab.full[(((max(numvar)/2)*3)+1):(((max(numvar)/2)*3)*2),],
                           align = c("llcccc")),
      include.rownames = FALSE,
      type             = "html")
```

## `r as.numeric(last(refyear))` Costs

The total annual cost for incarcerating **probation violators** in `r as.numeric(last(refyear))` is `r dollar_format()(total.pr.vp)`. The total annual cost for incarcerating **parole violators** in `r as.numeric(last(refyear))` is `r dollar_format()(total.pa.vp)`.

<br>

\noindent The total annual cost for incarcerating just those with **technical violations** in `r as.numeric(last(refyear))` is `r dollar_format()(total.tvp)`.

<br>

\noindent The total annual cost for incarcerating just those with **new offense violations** in `r as.numeric(last(refyear))` is `r dollar_format()(total.novp)`.

<br>

### `r as.numeric(last(refyear))` Costs by State

Below is a table of these costs by state (no imputation). Note that missing cells indicate that a state did not submit `r as.numeric(last(refyear))` population data:

<br>

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
print(xtable(cost.final[c("State","cost.vp.state","cost.tvp.state")] %>%
                              mutate(
                                `Community Supervision Violator Cost` = case_when(cost.vp.state >=0 ~ dollar_format()(cost.vp.state),
                                                                                  TRUE ~ NA_character_),
                                `Technical Violator Cost`             = case_when(cost.vp.state >=0 ~ dollar_format()(cost.tvp.state),
                                                                                  TRUE ~ NA_character_)
                                     ) %>%
                              select(-c(cost.vp.state,cost.tvp.state)),
                           align = c("llrr")),
      include.rownames = FALSE,
      type             = "html")
```