---
title: "2022 MCLC Survey Results"
author: "The CSG Justice Center"
date: "4/4/2023"
output:
  html_document
---

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
#######################################
# PROJECT: More Community, Less Confinement (2022)
# PURPOSE: Impute values, national estimates, state estimates and costs
# AUTHOR: JSM
#######################################

###################################################
#OVERALL SETUP
#load necessary packages
library(openxlsx)
library(dplyr)
library(readr)
library(reshape)
library(readxl)
library(tidyverse)
library(data.table)
library(formattable)
library(scales)
library(mice)
library(VIM)
library(finalfit)
library(janitor)
library(Hmisc)
library(eeptools)
library(csgjcr)
library(scales)
library(xtable)

# #name of data to import and path
# sp_data_path <- csgjcr::csg_sp_path("50 State Revocations Project/50 State Survey (2022)/Data")
# dataname     <- "/mclc_data_2022_v4.xlsx"

# #read automated_clean to get data, reformat
# source("02_clean.R")
source("03_clean.R")

##SETUP
#years of data
year     <- c('2018','2019','2020','2021')
yearnum  <- c(1,     2,     3,     4)
refyear  <- c('2019','2020','2021')
refnum   <- 1:3

#number of survey variables
numvar   <- 1:16

#labels
var.labels = c(year                                = "Year",
               overall_admissions                  = "Overall admissions",
               admissions_for_violations           = "Admissions for violations",
               admissions_for_technical_violations = "Admissions for technical violations",
               admissions_for_new_crime_violations = "Admissions for new crime violations",
               technical_probation_violation_admissions= "Admissions for technical violations, probation", 
               technical_parole_violation_admissions= "Admissions for technical violations, parole",
               new_offense_probation_violation_admissions= "Admissions for new crime violations, probation",
               new_offense_parole_violation_admissions= "Admissions for new crime violations, parole",
               
               overall_population                  = "Overall population",
               violator_population                 = "Violator population",
               technical_violator_population       = "Technical violator population",
               new_crime_violator_population       = "New crime violator population",
               technical_probation_violation_population= "Technical violator population, probation",
               technical_parole_violation_population= "Technical violator population, parole",
               new_offense_probation_violation_population= "New crime violator population, probation",
               new_offense_parole_violation_population= "New crime violator population, parole"
)

################################################################################
# IMPUTATION
# MICE
# National Estimates w/95% CI's
################################################################################

#calculate imputed values (multiple imputation)
source("MImp1.R")

######
#produce confidence intervals
#post multiple imputation t-test
#grab response columns for creating estimated counts/CIs
######
#get response columns from survey
#there are 8 response columns
tablevals<- colnames(mice_imputed_data1[numvar[3]:(length(numvar)+2)])

source("MImp2.R")

##############################################################
##################CREATE A TABLE HERE!!!!!!!!!!!!#############
#combine
table2018 <- cbind(nat.2018all,natCI.2018all) %>%
  mutate(across(where(is.numeric),formattable::comma,0),
         id     = 1:16,
         ' '    = '95% Confidence Interval',
         '2018' = paste0('(',`1`,', ',`2`,')')
         )
table2019 <- cbind(nat.2019all,natCI.2019all) %>%
  mutate(across(where(is.numeric),formattable::comma,0),
         '2019' = paste0('(',`1`,', ',`2`,')'))
table2020 <- cbind(nat.2020all,natCI.2020all) %>%
  mutate(across(where(is.numeric),formattable::comma,0),
         '2020' = paste0('(',`1`,', ',`2`,')'))
table2021 <- cbind(nat.2021all,natCI.2021all) %>%
  mutate(across(where(is.numeric),formattable::comma,0),
         '2021' = paste0('(',`1`,', ',`2`,')'))

tablevals.pretty <- c('Overall Admissions',
                     'Admissions for Supervision Violations',
                     'Admissions for Technical Supervision Violations',
                     'Admissions for New Crime Violations',
                     'Admissions for Technical Probation Violations',
                     'Admissions for Technical Parole Violations',
                     'Admissions for New Crime Probation Violations',
                     'Admissions for New Crime Parole Violations',
                     
                     'Overall Population',
                     'Population of Supervision Violators',
                     'Population of Technical Supervision Violators',
                     'Population of New Crime Violators',
                     'Population of Technical Probation Violations',
                     'Population of Technical Parole Violations',
                     'Population of New Crime Probation Violations',
                     'Population of New Crime Parole Violations'
                     )

tab.full <- data.table(cbind(table2018$Estimate,table2019$Estimate,table2020$Estimate,table2021$Estimate)) %>%
  mutate(id = 1:16,
         ' ' = 'Estimate', .before="V1") %>%
  dplyr::rename('2018' = "V1",
                '2019' = "V2",
                '2020' = "V3",
                '2021' = "V4") %>%
  rbind(
    data.table(' ' = tablevals.pretty,
               id  = 1:16,
               id2 = 0),
    
    data.table(
      cbind(
      ' '    = table2018$` `,
      '2018' = table2018$`2018`,
      '2019' = table2019$`2019`,
      '2020' = table2020$`2020`,
      '2021' = table2021$`2021`,
      'id'   = table2018$id
      )),
    fill = TRUE) %>%
  mutate(id = as.numeric(id)) %>%
  arrange(id,id2) %>%
  select(-c(id,id2))

saveRDS(tab.full, file = "C:/Users/jmallett/OneDrive - The Council of State Governments/GitProjects/MCLC_survey/code/survey/tables_lowagg.RDS")

##############################################################
##############################################################

#############################################################################
###############CALCULATING STATE ESTIMATES
#############################################################################

source("MImp3.R")

#############################
###COSTS#####################

source("Costs.R")

#averted cost total
avert19_20last <- 2742725774
avert19_20     <- sum(cost.final$averted_costs19_20, na.rm = TRUE) #2019 - 2020
avertdiff      <- avert19_20last - avert19_20 
avert19_21     <- sum(cost.final$averted_costs19_21, na.rm = TRUE) #2019 - 2021
avert20_21     <- sum(cost.final$averted_costs20_21, na.rm = TRUE) #2020 - 2021

##Annual cost for incarcerating community supervision violators
total.vp21 <- sum(cost.final$cost.vp21) #national

##Annual cost for incarcerating those w/just technical violations
total.tvp21 <- sum(cost.final$cost.tvp21) #national

```

## National Estimates

These estimates incorporate BJS total population and admissions data in calculations across all years.

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}

print(xtable(tab.full[1:24,],
                           align = c("llcccc")),
      include.rownames = FALSE,
      type             = "html")

```

<br>
<br>


```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
print(xtable(tab.full[25:48,],
                           align = c("llcccc")),
      include.rownames = FALSE,
      type             = "html")
```

## 2021 Costs

The total annual cost for incarcerating community supervision violators in 2021 is `r dollar_format()(total.vp21)`.

<br>

\noindent The total annual cost for incarcerating just those with technical violations in 2021 is `r dollar_format()(total.tvp21)`.

<br>

### 2021 Costs by State

Below is a table of these costs by state (no imputation). Note that missing cells indicate that a state did not submit 2021 population data:

<br>

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
print(xtable(cost.final[c("State","cost.vp21.state","cost.tvp21.state")] %>%
                              mutate(
                                `Community Supervision Violator Cost` = case_when(cost.vp21.state >=0 ~ dollar_format()(cost.vp21.state),
                                                                                  TRUE ~ NA_character_),
                                `Technical Violator Cost`             = case_when(cost.vp21.state >=0 ~ dollar_format()(cost.tvp21.state),
                                                                                  TRUE ~ NA_character_)
                                     ) %>%
                              select(-c(cost.vp21.state,cost.tvp21.state)),
                           align = c("llrr")),
      include.rownames = FALSE,
      type             = "html")
```


## Difference in National Estimates

The table below displays results of the difference in National estimates from the previous method (multiple imputation of 8 metrics) to the new method (multiple imputation of 16 metrics, including lowest level of aggregation).

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
#compare national estimates when including lowest aggregation in multiple imputation
new <- readRDS("tables_lowagg.RDS") 
new1 <- new %>% #feature branch
  mutate(
    id = case_when(` ` == "Overall Admissions" ~ 1,
                   ` ` == "Estimate" & row.names(new)=="2" ~ 2,
                   ` ` == "95% Confidence Interval" & row.names(new)=="3" ~ 3,
                   ` ` == "Admissions for Supervision Violations" ~ 4,
                   ` ` == "Estimate" & row.names(new)=="5" ~ 5,
                   ` ` == "95% Confidence Interval" & row.names(new)=="6" ~ 6,
                   ` ` == "Admissions for Technical Supervision Violations" ~ 7,
                   ` ` == "Estimate" & row.names(new)=="8" ~ 8,
                   ` ` == "95% Confidence Interval" & row.names(new)=="9" ~ 9,
                   ` ` == "Admissions for New Crime Violations" ~ 10,
                   ` ` == "Estimate" & row.names(new)=="11" ~ 11,
                   ` ` == "95% Confidence Interval" & row.names(new)=="12" ~ 12,
                   ` ` == "Admissions for Technical Probation Violations" ~ 13,
                   ` ` == "Estimate" & row.names(new)=="14" ~ 14,
                   ` ` == "95% Confidence Interval" & row.names(new)=="15" ~ 15,
                   ` ` == "Admissions for Technical Parole Violations" ~ 16,
                   ` ` == "Estimate" & row.names(new)=="17" ~ 17,
                   ` ` == "95% Confidence Interval" & row.names(new)=="18" ~ 18,
                   ` ` == "Admissions for New Crime Probation Violations" ~ 19,
                   ` ` == "Estimate" & row.names(new)=="20" ~ 20,
                   ` ` == "95% Confidence Interval" & row.names(new)=="21" ~ 21,
                   ` ` == "Admissions for New Crime Parole Violations" ~ 22,
                   ` ` == "Estimate" & row.names(new)=="23" ~ 23,
                   ` ` == "95% Confidence Interval" & row.names(new)=="24" ~ 24,
                   ` ` == "Overall Population" ~ 25,
                   ` ` == "Estimate" & row.names(new)=="26" ~ 26,
                   ` ` == "95% Confidence Interval" & row.names(new)=="27" ~ 27,
                   ` ` == "Population of Supervision Violators" ~ 28,
                   ` ` == "Estimate" & row.names(new)=="29" ~ 29,
                   ` ` == "95% Confidence Interval" & row.names(new)=="30" ~ 30,
                   ` ` == "Population of Technical Supervision Violators" ~ 31,
                   ` ` == "Estimate" & row.names(new)=="32" ~ 32,
                   ` ` == "95% Confidence Interval" & row.names(new)=="33" ~ 33,
                   ` ` == "Population of New Crime Violators" ~ 34,
                   ` ` == "Estimate" & row.names(new)=="35" ~ 35,
                   ` ` == "95% Confidence Interval" & row.names(new)=="36" ~ 36,
                   ` ` == "Population of Technical Probation Violations" ~ 37,
                   ` ` == "Estimate" & row.names(new)=="38" ~ 38,
                   ` ` == "95% Confidence Interval" & row.names(new)=="39" ~ 39,
                   ` ` == "Population of Technical Parole Violations" ~ 40,
                   ` ` == "Estimate" & row.names(new)=="41" ~ 41,
                   ` ` == "95% Confidence Interval" & row.names(new)=="42" ~ 42,
                   ` ` == "Population of New Crime Probation Violations" ~ 43,
                   ` ` == "Estimate" & row.names(new)=="44" ~ 44,
                   ` ` == "95% Confidence Interval" & row.names(new)=="45" ~ 45,
                   ` ` == "Population of New Crime Parole Violations" ~ 46,
                   ` ` == "Estimate" & row.names(new)=="47" ~ 47,
                   ` ` == "95% Confidence Interval" & row.names(new)=="48" ~ 48)
  )
    
orig <- readRDS("tables_orig8metrics.RDS")
orig1 <- orig %>% #main branch
  mutate(
    id = case_when(` ` == "Overall Admissions" ~ 1,
                   ` ` == "Estimate" & row.names(orig)=="2" ~ 2,
                   ` ` == "95% Confidence Interval" & row.names(orig)=="3" ~ 3,
                   ` ` == "Admissions for Supervision Violations" ~ 4,
                   ` ` == "Estimate" & row.names(orig)=="5" ~ 5,
                   ` ` == "95% Confidence Interval" & row.names(orig)=="6" ~ 6,
                   ` ` == "Admissions for Technical Supervision Violations" ~ 7,
                   ` ` == "Estimate" & row.names(orig)=="8" ~ 8,
                   ` ` == "95% Confidence Interval" & row.names(orig)=="9" ~ 9,
                   ` ` == "Admissions for New Crime Violations" ~ 10,
                   ` ` == "Estimate" & row.names(orig)=="11" ~ 11,
                   ` ` == "95% Confidence Interval" & row.names(orig)=="12" ~ 12,

                   ` ` == "Overall Population" ~ 25,
                   ` ` == "Estimate" & row.names(orig)=="14" ~ 26,
                   ` ` == "95% Confidence Interval" & row.names(orig)=="15" ~ 27,
                   ` ` == "Population of Supervision Violators" ~ 28,
                   ` ` == "Estimate" & row.names(orig)=="17" ~ 29,
                   ` ` == "95% Confidence Interval" & row.names(orig)=="18" ~ 30,
                   ` ` == "Population of Technical Supervision Violators" ~ 31,
                   ` ` == "Estimate" & row.names(orig)=="20" ~ 32,
                   ` ` == "95% Confidence Interval" & row.names(orig)=="21" ~ 33,
                   ` ` == "Population of New Crime Violators" ~ 34,
                   ` ` == "Estimate" & row.names(orig)=="23" ~ 35,
                   ` ` == "95% Confidence Interval" & row.names(orig)=="24" ~ 36)
  )


#compare difference in multiple imputation using different approaches (just estimates)
compare <- merge(orig1,new1,by = "id") %>%
  filter(` .x` != "95% Confidence Interval") %>%
  mutate(
    `Difference in 2018` = as.numeric(`2018.x`) - as.numeric(`2018.y`),
    `Difference in 2019` = as.numeric(`2019.x`) - as.numeric(`2019.y`),
    `Difference in 2020` = as.numeric(`2020.x`) - as.numeric(`2020.y`),
    `Difference in 2021` = as.numeric(`2021.x`) - as.numeric(`2021.y`)
  ) %>%
  select(c(` .x`, 
           `Difference in 2018`, 
           `Difference in 2019`, 
           `Difference in 2020`, 
           `Difference in 2021`)) %>%
  dplyr::rename(` ` = ` .x`) %>%
  mutate(
    `Difference in 2018` = ifelse(!is.na(`Difference in 2018`),format(as.numeric(`Difference in 2018`), nsmall=0, digits = 3, scientific = F,big.mark=","),''),
    `Difference in 2019` = ifelse(!is.na(`Difference in 2019`),format(as.numeric(`Difference in 2019`), nsmall=0, digits = 3, scientific = F,big.mark=","),''),
    `Difference in 2020` = ifelse(!is.na(`Difference in 2020`),format(as.numeric(`Difference in 2020`), nsmall=0, digits = 3, scientific = F,big.mark=","),''),
    `Difference in 2021` = ifelse(!is.na(`Difference in 2021`),format(as.numeric(`Difference in 2021`), nsmall=0, digits = 3, scientific = F,big.mark=","),'')
  )

#print table
print(xtable(compare[1:16,],
                           align = c("llcccc")),
      include.rownames = FALSE,
      type             = "html")

```
